---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```
# sneezy

Explaining t-SNE via diagnostics and tours.

# Installation

```{r install, eval = FALSE}
remotes::install_github("sa-lee/sneezy")
```

# Quick start

sneezy comes preloaded with some datasets from the 
[multi-challenge dataset](http://ifs.tuwien.ac.at/dm/dataSets.html). Let's 
look at a example, consisting two well sepearted Gaussian clusters
in 10 dimensional space (with same covariance matrix), and two 3-dimensional
Gaussian clusters embeded in 10 dimensional space. One of the clusters has more granular
structure and consists of three sub clusters. From the multichallenge dataset
page this subset is described as follows

> The first subset consists of a Gaussian cluster and another cluster that is itself divided into three Gaussian clusters, all of them living in a three-dimensional space. This subset is used to demonstrate how an algorithm deals with different levels of granularity in cluster structures. The distance between the centers of the two main clusters, i.e. the the big cluster and the cluster that consists of the three smaller ones, is 5 times the standard deviation d of the first main cluster. The three smaller clusters are arranged around the center of the second cluster, which they themselves form, on a circle of radius 5-d equidistant from each other. The three 3 small cluster centers and the center of the large cluster lie in the same plane. The small clusters each have a standard deviation of d and one third of the number of 3 data points of the large cluster.

## Making shapes

We can get a view of the structure using principal components:

```{r}
library(ggplot2)
library(gganimate) # required for printing tours
library(sneezy)
spheres <- subset(multi, key %in% c("A", "D"))
labels <- ifelse(spheres$key == "A", "sub-gaussian", "10-d 2x cluster")
spheres <- as.matrix(spheres[, -c(1,2)])

pc <- prcomp(spheres)
asp <- sqrt(pc$sdev[1] / pc$sdev[2])
ggplot(as.data.frame(pc$x), aes(PC1, PC2)) +
  geom_point(aes(colour = labels)) +
  coord_fixed(asp)

```

And the equivalent t-SNE, with our simplified wrapper which computes
exact t-SNE for a given perplexity and exaggeration factor alpha. 

In this case it looks as though t-SNE has worked well: it has identified 
the three  subclusters of the second cluster embededded in 3-dimensions, 
and seperated the two 10-d clusters.

```{r}
set.seed(1010010)
coords <- basic_tsne(spheres, perplexity = 30)
ggplot(as.data.frame(coords$Y), aes(V1, V2)) +
  geom_point(aes(colour = labels)) +
  coord_fixed(asp) +
  scale_color_brewer(palette = "Dark2")
```


We can also tour around a data space, and see how the nearest neighbours
graph from t-SNE space is preserved in high-dimensional space. We can take
subsets of the nn graph to see how the t-SNE preserves local topology, 
for example the points on the outside of the subcluster on the right hand side:


```{r}
pal <- c("#1B9E77", "#D95F02")[as.integer(as.factor(labels))]
sneezy_neighbours(spheres, coords, .subset = 171, col = pal)
```

We can also triangulate the points in t-SNE space

```{r}
sneezy_triangles(spheres, coords, col = pal)
```

And look at the centroids in the original space of the nearest neighbours
graph in t-SNE space:

```{r}
sneezy_centroids(spheres, coords)
```
