---
title: "Introducing sneezy"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{sneezy}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, include = FALSE}
library(sneezy)
library(SingleCellExperiment)
# to display gifs inline
library(gganimate)
```

# Exploring non-linear embeddings with sneezy

The goal of sneezy is to allow a user to exlpore non-linear embedding
algorithms such as t-SNE and UMAP in the context of global geometry of
the original high dimensional space. It does this in a few ways:

* generating a tour of the original (high dimensional) data set
* estimating and visualising nearest neighbor graphs in the original and embedding space
* providing numerical and visual diagnostics to determine the quality of the embedding

It provides a unified interface to a selection of embedding methods 
and nearest neighbor estimation methods. 


# The multi-challenge dataset

sneezy comes preloaded with a benchmark from the 
[multi-challenge dataset](http://ifs.tuwien.ac.at/dm/dataSets.html). 

This data arrives as a simple `data.frame` that contains a key indicating 
the challenge id and an index giving the observation id for that particular
challenge, the remaining coordinates are the dimensions.

```{r}
library(SingleCellExperiment)
library(sneezy)
head(multi)
```

## The TourExperiment object

In order to keep track of the embeddings and transformations on the
numerical parts of the data, sneezy creates a data structure called a
TourExperiment. It extends the `SingleCellExperiment::SingleCellExperiment()`
object with two additional slots called `basisSets` and `neighborSets`,
we will describe their use in more detail shortly. 

A TourExperiment object can be created directly from a data.frame or a matrix as follows:

```{r}
multi_te <- TourExperiment(multi,
                           basisSets = SimpleList(),
                           neighborSets = SimpleList(),
                           X1:X10)
multi_te
```

The last argument is the selection of columns that form the numeric data
that we are interested in exploring the structure of, called an assay
in the nonmelcature of bioinformatics. Note this is transposed so rows are
variables and columns are observations.

```{r}
# the x1 - x10 data
assay(multi_te, "view")[, 1:5]
```

## Making Shapes

We will start with a simpler look of the data by selecting the keys "A"
and "D". This consists of two well sepearted Gaussian clusters
in 10 dimensional space (with same covariance matrix), and two 3-dimensional
Gaussian clusters embeded in 10 dimensional space. One of the clusters has more granular
structure and consists of three sub clusters. From the multichallenge dataset
page this subset is described as follows:

> The first subset consists of a Gaussian cluster and another cluster that is itself divided into three Gaussian clusters, all of them living in a three-dimensional space. This subset is used to demonstrate how an algorithm deals with different levels of granularity in cluster structures. The distance between the centers of the two main clusters, i.e. the the big cluster and the cluster that consists of the three smaller ones, is 5 times the standard deviation d of the first main cluster. The three smaller clusters are arranged around the center of the second cluster, which they themselves form, on a circle of radius 5-d equidistant from each other. The three 3 small cluster centers and the center of the large cluster lie in the same plane. The small clusters each have a standard deviation of d and one third of the number of 3 data points of the large cluster.

First we will subset the `multi_te` object:

```{r}
multi_te2 <- multi_te[, colData(multi_te)$key %in% c("A", "D")]
colData(multi_te2)$labels <- ifelse(colData(multi_te2)$key == "A", 
                                    "sub-gaussian", 
                                    "10-d 2x cluster")
multi_te2
```

We can get a view of the structure using exact principal components via
the `embed_linear` function:

```{r}
multi_te2 <- embed_linear(multi_te2, 
                          num_comp = nrow(multi_te2),
                          .on = "view",
                          center = TRUE,
                          .engine = pca_exact())
multi_te2
```

This adds an element to the reducedDims slot called `pca_exact()` that
stores the sample factors, and loadings for the principal components.

```{r}
pcs <- reducedDim(multi_te2, "pca_exact")
pcs
```

The resulting principal components can be viewed via ggplot2:

```{r}
library(ggplot2)
ggplot(
  data.frame(sampleFactors(pcs), labels = colData(multi_te2)$labels),
  aes(x = PC1, y = PC2)
  ) +
  geom_point(aes(colour = labels)) +
  scale_color_brewer(palette = "Dark2")
```

From the PCs we can clearly see the separation of the two 10-d clusters
but cannot see the finer structure in the sub-gaussian clusters. 

At this point we can compute exact t-SNE for a given perpelexity and
exaggeration factor alpha:

```{r}
set.seed(1010010)
multi_te2 <- embed_nonlinear(
  multi_te2,
  num_comp = 2,
  .on = "view", 
  .engine = tsne_exact(perplexity = 30, alpha = ncol(multi_te2) / 10)
)

tsne <- reducedDim(multi_te2, "tsne_exact")
```

And we can follow the same approach as before for 
visualising the results:

```{r}
pl <- ggplot(
  data.frame(sampleFactors(tsne), labels = colData(multi_te2)$labels),
  aes(x = Dim1, y = Dim2)
) +
  geom_point(aes(colour = labels)) +
  scale_color_brewer(palette = "Dark2") +
  theme(axis.title = element_blank(), 
        axis.ticks = element_blank(), 
        axis.text = element_blank(),
        panel.grid = element_blank())

pl  

```

In this case it looks as though t-SNE has worked ok: it has identified 
the three subclusters of the second cluster for the orange points, 
and seperated the two ten dimensional clusters.

Next we can estimate the nearest neighbors on both the original dataset
and in the embedding. This produces a matrix of indices for each row
giving the nearest neighbor. 

```{r}
multi_te2 <- estimate_neighbors(multi_te2, 90, .on = "view")
multi_te2 <- estimate_neighbors(multi_te2, 90, .on = "tsne_exact")

dim(neighborSet(multi_te2, "tsne_exact"))
```

You can then view the k-nearest neighbors graph overlaid on the
scatter via the `overlay_neighbors()` function:

```{r}
pl +
  overlay_neighbors(sampleFactors(tsne)[,1], 
                    sampleFactors(tsne)[,2], 
                    neighborSet(multi_te2, "tsne_exact"),
                    alpha = 0.01)

```

This can give us a direct visualisation of how the difference
between the emphasis t-SNE provides on local geometry by viewing
the k-NN graph from the original dataset on the t-SNE layout.

```{r}
pl +
  overlay_neighbors(sampleFactors(tsne)[,1], 
                    sampleFactors(tsne)[,2], 
                    neighborSet(multi_te2, "view"),
                    alpha = 0.01)
```

We can directly visualise the difference: 

```{r}
pl +
  overlay_difference_neighbors(sampleFactors(tsne)[,1], 
                    sampleFactors(tsne)[,2], 
                    neighborSet(multi_te2, "view"),
                    neighborSet(multi_te2, "tsne_exact"),
                    alpha = 0.01)
```

Or the intersection:

```{r}
pl +
  overlay_intersect_neighbors(sampleFactors(tsne)[,1], 
                    sampleFactors(tsne)[,2], 
                    neighborSet(multi_te2, "view"),
                    neighborSet(multi_te2, "tsne_exact"),
                    alpha = 0.01)
```


Sometimes the graph view can be messy, especially when there's lots of neighbors. Instead we can estimate the centroids of clusters seen in the
embedding by clustering the complete kNN or the shared nearest neighbor (sNN) graph. Note that this is heavily influenced by the use of  

```{r}
pl +
  overlay_snn_centroids(sampleFactors(tsne)[,1], 
                        sampleFactors(tsne)[,2], 
                        neighborSet(multi_te2, "tsne_exact"),
                        color = "red") 
```

## Viewing high-dimensional space

We can also tour around a data space, and see how the nearest neighbours
graph from t-SNE space is preserved in high-dimensional space. We can take view the entire k-NN or s-NN graph in the embedding space.
This allows us to see how t-SNE preserves local topology. First
we generate a basis set on the  for our tour, which will form part of
our `TourExperiment`:


```{r}
set.seed(1999)
multi_te2 <- generate_bases(multi_te2, .on = "view", max_bases = 300)
basisSetNames(multi_te2)
```

Then we can animate the tour view and overlay the neighbors
with the following:

```{r}
library(gganimate)
pal <- c("#1B9E77","#D95F02")[as.integer(factor(colData(multi_te2)$labels))]

sneezy_neighbors(multi_te2, "view", "tsne_exact", col = pal)
```

And look at the centroids in the original space of the nearest neighbours graph in t-SNE space:

```{r}
sneezy_centroids(multi_te2, "view", "tsne_exact", col = pal)
```
